---
title: "Sta 323 - Midterm 2 - Spring 2022"
author: "Aidan Gildea"
output: rmarkdown::html_document
runtime: shiny
---
Due Wednesday, April 20th by 5:00 pm

<br/>

```{r lib}
library(httr)
library(rvest)
library(tidyr)
library(dplyr)
library(repurrrsive)
library(validate)
library(tibble)
```



### Rules

Review all of the rules detailed in `README.md`, if you have any questions please direct them to myself or the TAs.

<br/>

### Task 1 - Figuring out the NY Times Article Search API

``` {r task 1}
bday_url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json?fq=document_type:article AND print_page:1 AND print_section:A&begin_date=20191115&end_date=20191115&api-key=V7UTIeAubbaAzsibQRlIJ27QE2wweTVz"
```

* `document_type=` article

* `print_page=` 1

* `print_section=` A

* `fq=` document_type:article AND print_page:1 AND print_section:A

* `begin_date=` 20191115

* `end_date=` 20191115


<br/>    

### Task 2 - Getting data from the NY Times Article Search API

To access the appropriate articles from a particular day of interest, we
can create a get_nyt_articles() function with four parameters: year, month, day, and api_key. To ensure these inputs will function when used in our request to the New York Times API, we begin with a few basic sanity checks. These checks make sure that 1) month is a number between 1 and 12 with length 1, 2) day is a number between 1 and 31 with length 1, and 3) year is a four digit number with length 1. If the inputs pass these conditions, then we can create the individual parts of our NYT API request URL. This includes our base url (which specifies article type, print page 1, and print section A), our begin and end date (in the format YYYYMMDD), and then our api-key. Now we can begin our request to the NYT API. Our request must account for multiple pages of articles (in the case there are more than 10 articles), therefore, we will use a repeat loop that deals with pagination. Beginning at page 0, we will make a request to the API using our URL and then check if the number of articles is less than our page size of 10. If it is more than 10, we will add the articles, increase the page by 1, and repeat the loop; otherwise, we will add the articles and end the loop. To account for the rate limit and timeouts, we suspend execution for 6 seconds each loop - as recommended in the README. If the function determines no articles were found for the specified date, it will cast an error message and return an empty dataframe. Otherwise, it handles the list of articles, using unnest() and hoist() calls to return a final tidy dataframe of the results.

```{r task 2, message=FALSE}
get_nyt_articles = function(year, month, day, api_key) {
  
  # validate user's input for month (value of length 1, between [1, 12])
  if (length(month) != 1 || month < 1 || month > 12)
  {
    validate("Month input must be a number in the range of [1, 12]")
  }     
  # validate user's input for day (value of length 1, between [1, 31])
  if (length(day) != 1 || day < 1 || day > 31)
  {
    validate("Day input must be a number in the range of [1, 31]")
  }     
  # validate user's input for year (value is numeric, length of 1, has 4 digits)
  if (!is.numeric(year) || length(year) != 1 || nchar(year) != 4) {
    validate("Year input must be a four-digit number")
  }
  # add leading zero to months less than 10
  if (month < 10) {
    month = paste0("0", month)
  }
  # add leading zero to days less than 10
  if (day < 10) {
    day = paste0("0", day)
  }
  # create base url
  base_url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json?fq=document_type:article AND print_page:1 AND print_section:A&"
  # specify begin date
  begin_date <- paste0("begin_date=", year, month, day, "&")
  # specify end date
  end_date <- paste0("end_date=", year, month, day, "&")
  # save user api-key
  key <- paste0("api-key=", api_key)
  # create empty list to store results of pagination
  all_articles=list()
  # set page initially to 0
  page = 0
  # specify NYT API request limit
  pageSize = 10
  # create empty dataframe
  nyt_df = data.frame() %>%
    add_column(abstract = character(0), web_url = character(0), snippet = character(0), lead_paragraph= character(0), print_section= character(0), print_page= character(0), source= character(0), main_headline= character(0), pub_date= character(0), document_type= character(0), news_desk= character(0), section_name= character(0), main_byline= character(0), type_of_material= character(0), word_count= character(0)) %>%
    as_tibble()
  # create function that loops through pages of particular api request
  repeat{
    cat("Downloading page", page, "\n")
    # retrieve information from url based on user inputs
    url = jsonlite::read_json(URLencode(glue::glue(base_url, "page:{page}&", begin_date, end_date, key)))
    # retrieve articles from request
    art_list <-  url$response$docs
    # check if request is below NYT APT document limit
    if (length(art_list) < pageSize){
      # add to article list
      all_articles = c(all_articles, art_list)
      # break from repeat loop is less than 10 documents
      break
    }
    # account for timeouts
    Sys.sleep(6) 
    # adds next iteration of articles to full list
    all_articles = c(all_articles, art_list)
    # advances to next page
    page = page+1
  }
  # checks if no articles were retrieved
  if (length(all_articles) == 0) {
    # casts appropriate warning
    warning("There are no results on the specified date.")
    # returns an empty dataframe
    return(nyt_df)
  }
  # otherwise, articles were present on specified date
  else{
    # create tibble
    nyt_df <- (nyt_df = tibble::tibble(
      info = all_articles
    ))
    # handle tibble to retrieve variables of interest
    nyt_tidy <- nyt_df %>%
      unnest_wider(info) %>%
      hoist(headline,
            main_headline = "main", .remove = TRUE) %>%
      hoist(byline,
            main_byline = "original") %>%
      # retrieve extranneous extra data
      filter(!is.na(web_url))%>%
      # select variables of interest
      select(abstract, web_url, snippet, lead_paragraph, print_section, print_page, source, main_headline, pub_date, document_type, news_desk, section_name, main_byline, type_of_material, word_count)
  }
  # return tidy data frame
  return(nyt_tidy)
}
```


<br/>

### Task 3 - Shiny Front End

This New York Times API shiny app features a side panel where users can input their year, month, and day of interest. Then, after inputting their API key and pressing the search action button, the app will produce a list of headlines from articles on the specified date. The shiny app does so by calling the get_nyt_articles() function created in task 2 when the search action button is pressed. If there are no articles present on that day, the app will output a message indicating so. Otherwise, the server uses a map to create a dynamic list of links labeled by their respective headlines. This dynamic list of actionLinks allows users to then click on the links, prompting a modal dialog box featuring the article's headline, byline, first paragraph, word count, and hyperlink. These modal dialog boxes are conditional on users clicking on the respective headlines (as seen by the use of an observeEvent). By using an uiOutput in the main panel, the app is dynamic and will update its list of articles each time a user changes their inputs and presses the search button. 

```{r}
# load packages
library(shiny)
library(shinythemes)
library(purrr)
# create shiny app
shinyApp(
  # implement user interface
  # add theme for aesthetic purposes
  ui = fluidPage(theme = shinytheme("lumen"),
                 # name app
                 titlePanel("NYT Article Database"),
                 # create side bar panel
                 sidebarLayout(
                   sidebarPanel(
                     # user input for year (max updates based on current year)
                     numericInput("year","Year", value = 2019, min = 1900, max = format(Sys.Date(), "%Y")),
                     # user input for month (min of 1, max of 12)
                     numericInput("month","Month", value = 11, min = 1, 
                                  max = 12),
                     # user input for day (min of 1, max of 31)
                     numericInput("day","Day", value = 15, min = 1, max = 31),
                     # user input for key
                     textInput("key", "Enter API Key:", value = "V7UTIeAubbaAzsibQRlIJ27QE2wweTVz"),
                     # search button that produces corresponding inputs
                     h4("Retrieve Articles:"),
                     actionButton("run", "Search"),
                   ),
                   mainPanel(
                     # header in main panel
                     h3(HTML("<u>Articles<u>")),
                     # dynamic output of links according to selected date
                     uiOutput("links")
                   )
                 )
  ),
  server = function(input, output, session) {
    state = reactiveValues(
      observers = list()
    )
    
    # event dependent on user pressing the search button
    observeEvent(input$run, {
      # retrieve articles according to selected date
      articles = get_nyt_articles(input$year, input$month, input$day, input$key)
      # checks to see if there are no article on corresponding date
      if(is_empty(articles$web_url)) {
        # outputs message indicating there are no articles
        output$links = renderUI(
          fluidPage("There are no articles on the specified date."))
        # otherwise, there are articles corresponding to the current date
      } else {
        # retrieve headlines from articles
        headlines <- articles$main_headline
        
        # Destroy existing observers
        for(i in seq_along(state$observers)) {
          state$observers[[i]]$destroy()
        }
        # dynamically create list of articles   
        ui_elems = map(
          # sequence length of amount of articles
          seq_len(length(headlines)), 
          # create action links per each article
          function(i) 
            fluidRow(actionLink(paste0("link",i), paste0("• ", headlines[i])))
        )
        # output the links created
        output$links = renderUI(fluidPage(ui_elems))
        
        # Reset and create new observers for each of our links
        state$observers = map(
          # sequence length of amount of articles
          seq_len(length(headlines)), 
          # create label for each link
          function(i) {
            label = paste0("link",i)
            # event dependent on user selecting particular link
            observeEvent(input[[label]], ignoreInit = TRUE, {
              # modal dialog pops up according the selected link
              showModal(modalDialog(
                # articles headline as title, byline as subtitle
                title = HTML(articles$main_headline[i], "<br>", articles$main_byline[i]),
                # base includes first paragraph of articles, word count, and                         link to full article
                HTML(articles$lead_paragraph[i], "<br><br>", "Word Count: ", articles$word_count[i], "<br><br>", "Read the full article at ", paste(a(" here.", href= articles$web_url[i]))),
                # button to close modal dialog
                footer = modalButton("Close")
              ))
            })
          }
        )
      }
    })
  }
)
```
